---
title: "Phagocytosis Analysis"
author: "Timothy Smyth"
date: "2024-03-21"
output: html_document
---

# Analysis of S. aureus bioparticle phagocysosis assay

### Load packages

```{r message = FALSE}
rm(list = ls(all.names = TRUE)) # clears global environ.

# Load packages
library(tidyverse) # for data cleaning
library(dplyr) # for data set manipulations
library(stringr)
library(ggrepel) # for pca text labeling
library(ggplot2) # for plotting
library(ggtext)
library(patchwork)
library(cowplot)
```

### Import and prepare data

```{r}
# Import data
Phagocytosis <- read.csv("Phagocytosis.csv")

colnames(Phagocytosis)[1] <- 'Particle_month'

Phagocytosis$Full_name <- paste0(Phagocytosis$Sample.ID, 
                                 "_", 
                                 Phagocytosis$Particle_month)

Phagocytosis <- data.frame(Phagocytosis, 
                           row.names = Phagocytosis$Full_name) %>% 
  subset(select = -c(Full_name))

# Rename particle months to full month name
Phagocytosis$Particle_month[Phagocytosis$Particle_month == 'Sept'] <- 'September'
Phagocytosis$Particle_month[Phagocytosis$Particle_month == 'Dec'] <- 'December'

# Retain a raw version of the phagocytosis data for fold change analysis below
Phagocytosis_Raw <- Phagocytosis

# Pivot data to long format
Phagocytosis <- Phagocytosis %>% 
  pivot_longer(cols = colnames(Phagocytosis[3:ncol(Phagocytosis)]), 
               names_to = 'Treatment', 
               values_to = 'MFI')

Phagocytosis$Sample.ID <- paste0(Phagocytosis$Sample.ID, 
                                 '_', 
                                 Phagocytosis$Particle_month)
                          
Phagocytosis$Treatment <- factor(Phagocytosis$Treatment, 
                                 levels = c('M0_Veh', 'M2_Veh', 
                                            'M0_M1', 'M2_M1', 
                                            'M0_PM', 'M2_PM', 
                                            'M0_M1.PM', 'M2_M1.PM'))
                       
Phagocytosis$Particle_month <- factor(Phagocytosis$Particle_month, 
                                      levels = c('December', 'March', 'June', 'September'))
```

### Calculate statistics and graph results for Supplemental Figure S1A

```{r}
# Isolate Vehicle exposed samples for Supplemental Figures S1A and S1B
Vehicle <- Phagocytosis %>% 
  subset(Treatment == 'M0_Veh' |
           Treatment == 'M2_Veh' |
           Treatment == 'M0_M1' |
           Treatment == 'M2_M1')

Vehicle$Treatment <- factor(Vehicle$Treatment, 
                            levels = c('M0_Veh', 'M2_Veh', 
                                       'M0_M1', 'M2_M1'))

# Create function to calculate statistics
statistics <- function(x){
  
  tmp <- x %>% 
    arrange(Treatment) %>%    
    mutate(Particle_month = factor(Particle_month,
                                   levels = c('December', 'March', 'June', 'September')))
  
  # Run Shapiro-Wilk normality test 
  # Testing each treatment group to determine if each falls within
  # a normal distribution for downstream statistical testing
  Shapiro <- Vehicle %>%
    group_by(Treatment) %>% 
    do(broom::tidy(shapiro.test(.$MFI)))
  
  # If the treatment group has a normal distribution
  # perform a paired t-test with BH correction
  if(min(Shapiro$p.value) >= 0.05){
    
    # Run pairwise t-test with BH correction
    stats <- rstatix::pairwise_t_test(MFI ~ Treatment, 
                                      paired = TRUE, 
                                      p.adjust.method = "BH", 
                                      data = x)
    
    # Round resulting p-values to 3 digits
    stats$adj.p.value <- round(stats$p.adj, 3)
    
    # Convert adjusted p values to significance symbols
    stats <- stats %>% mutate(adj.p.value = 
                                case_when((adj.p.value >= 0.05) ~ NA,
                                          
                                          (adj.p.value < 0.05 &
                                             adj.p.value >= 0.01) ~ '*', 
                                          
                                          (adj.p.value < 0.01 &
                                             adj.p.value >= 0.001) ~ '**',
                                          
                                          (adj.p.value < 0.001) ~ '***'))
    
    stats$test <- 't.test'
    
    stats
    
    } else{ # If treatment group does not have a normal distribution
      
      # Run a Wilcox test with BH correction
      stats <- rstatix::pairwise_wilcox_test(MFI ~ Treatment, 
                                    paired = TRUE, 
                                    p.adjust.method = "BH", 
                                    data = x)
      
      # Round resulting p-values to 3 digits
      stats$adj.p.value <- round(stats$p.adj, 3)
      
      # Convert adjusted p values to significance symbols
      stats <- stats %>% mutate(adj.p.value =
                                  case_when((adj.p.value >= 0.05) ~ NA,
                                            
                                            (adj.p.value < 0.05 &
                                               adj.p.value >= 0.01) ~ '*',
                                            
                                            (adj.p.value < 0.01 &
                                               adj.p.value >= 0.001) ~ '**',
                                            
                                            (adj.p.value < 0.001) ~ '***'))
      
      stats$test <- 'Wilcox'
      
      stats
      
    }
  
  if(all(is.na(stats[, 'adj.p.value'])) == TRUE){
      
      stats$pos_x_1 <- 0
      stats$pos_x_2 <- 0
      stats$pos_x <- 0
      stats$pos_y <- 0
      stats
      
      } else{
        
        stats <- na.omit(stats)
        
        stats <- stats %>% mutate(pos_x_1 =
                                    case_when((group1 == 'M0_Veh' | group1 == 'M0_PM') ~ 1,
                                              (group1 == 'M2_Veh' | group1 == 'M2_PM') ~ 2,
                                              (group1 == 'M0_M1' | group1 == 'M0_M1.PM') ~ 3, 
                                              (group1 == 'M2_M1' | group1 == 'M2_M1.PM') ~ 4))
        
        stats <- stats %>% mutate(pos_x_2 =
                                    case_when((group2 == 'M0_Veh' | group2 == 'M0_PM') ~ 1,
                                              (group2 == 'M2_Veh' | group2 == 'M2_PM') ~ 2,
                                              (group2 == 'M0_M1' | group2 == 'M0_M1.PM') ~ 3, 
                                              (group2 == 'M2_M1' | group2 == 'M2_M1.PM') ~ 4))
        
        stats$pos_x <- (stats$pos_x_1 + stats$pos_x_2)/2
        
        stats$pos_y <- seq(from = 0.25,
                           to = 0.25 * nrow(stats), 
                           0.25)
        
        stats
      }
    }

# Calculate statistics
stats <- statistics(Vehicle)

###########

# Determine the maximum data value for placement of significance bars
max <- max(Vehicle$MFI) * 1.05

# Define the locations for placement of significance bars
pos <- c(max, max * 1.15, max * 1.3, max * 1.45, max * 1.6, max * 1.75 )

# Generate Supplemental Figure S1A
Supplemental_Figure_S1A <- ggplot(data = Vehicle, 
                                  aes(x = Treatment, 
                                      y = MFI)) + 
  
  # Create a bar for each treatment group mean
  geom_bar(aes(col = Treatment), 
           position = position_dodge(),
           fill = 'white', 
           stat = "summary", 
           fun = "mean", 
           show.legend = FALSE) +
  
  # Create individual points for each data point
  # Set data point shapes to shapes indicated above
  geom_jitter(aes(col = Treatment),
              width = 0.25,
              size = 4,
              show.legend = FALSE) + 
  
  # Add SEM bars
  stat_summary(aes(col = Treatment),
               geom = "errorbar", 
               fun.data = mean_se, 
               position = "dodge") +
  
  xlab("Treatment") +
  
  ylab("Vehicle S. aureus \n Bioparticle Phagocytosis \n (MFI)") +
  
  labs(fill = "Particle Month") +
  
  theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 26),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 30), 
        
        plot.title = element_text(face = 'bold', 
                                  size = 20),
        
        legend.text = element_text(size = 20),
        
        legend.title = element_text(size = 20),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  
  scale_color_manual(name = 'Treatment', 
                     breaks = c('M0_Veh', 'M2_Veh', 'M0_M1', 'M2_M1'),
                     values = c('M0_Veh' = 'blue', 'M2_Veh' = 'red', 
                                'M0_M1' = 'darkgreen', 'M2_M1' = 'purple'))

###########
      
# Generate Supplemental Figure S1A statistics bars
Supplemental_Figure_S1A_Statistics <- ggplot(data = stats, 
                                             aes(x = group1,
                                                 y = pos_y)) +
  
  geom_segment(data = na.omit(stats), 
               aes(x = group1, 
                   xend = group2,
                   y = pos_y, 
                   yend = pos_y), 
               colour = "black", 
               linewidth = 1.25) +
  
  geom_text(data = stats,
            label = stats$adj.p.value,
            x = stats$pos_x,
            y = stats$pos_y,
            size = 14,
            colour = "black") +
  
  ylab("Vehicle S. aureus \n Bioparticle Phagocytosis \n (MFI)") +
  
  labs(fill = "Particle Month") +
  
  theme(axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.text.x = element_blank(),

        axis.title.x = element_blank(),

        plot.margin = unit(c(0, 0, 0, 0), "cm")) +

  scale_x_discrete(limits = c('M0_Veh', 'M2_Veh', 'M0_M1', 'M2_M1')) +
  scale_y_continuous(limits = c(0, max(stats$pos_y) + 0.25))

##########

# Combine Supplemental Figure S1A data and statistics plots
tiff('Supplemental_Figure_S1A.tiff', height = 1000, width = 1500)

wrap_plots(Supplemental_Figure_S1A_Statistics +  
             theme_void(),
           Supplemental_Figure_S1A + 
             theme(legend.position = "none"), 
           ncol = 1,
           heights = c(0.4, 2)) %>% 
  ggdraw()

dev.off()
```

### Calculate statistics and graph results for Supplemental Figure S1B

```{r}
Month_stats <- c('M0_Veh', 'M2_Veh', 'M0_M1', 'M2_M1')

# Create function to calculate statistics
Month_statistics <- function(x){
  
  tmp <- x %>% 
    arrange(Treatment) %>%    
    mutate(Particle_month = factor(Particle_month,
                                   levels = c('December', 'March', 'June', 'September')))
  
  # Run Shapiro-Wilk normality test 
  # Testing each particle month within a single treatment group to determine 
  # if each falls within a normal distribution for downstream statistical testing  
  Shapiro <- x %>%
    group_by(Particle_month) %>% 
    do(broom::tidy(shapiro.test(.$MFI)))
  
  # If the treatment group has a normal distribution
  # perform a paired t-test with BH correction
  if(min(Shapiro$p.value) >= 0.05){
    
    # Run pairwise t-test with BH correction
    stats <- rstatix::pairwise_t_test(MFI ~ Particle_month, 
                                      paired = FALSE, 
                                      p.adjust.method = "BH", 
                                      data = x)
    
    # Round resulting p-values to 3 digits
    stats$adj.p.value <- round(stats$p.adj, 3)
    
    # Convert adjusted p values to significance symbols
    stats <- stats %>% mutate(adj.p.value = 
                                case_when((adj.p.value >= 0.05) ~ NA,
                                          
                                          (adj.p.value < 0.05 &
                                             adj.p.value >= 0.01) ~ '*', 
                                          
                                          (adj.p.value < 0.01 &
                                             adj.p.value >= 0.001) ~ '**',
                                          
                                          (adj.p.value < 0.001) ~ '***'))
    
    
    stats$test <- 't.test'
    stats <- stats[, -c(7)] # Remove p.sigif to retain only adj.p.value
    stats
           
  } else{ # If treatment group does not have a normal distribution
    
    # Run a Wilcox test with BH correction
    stats <- rstatix::pairwise_wilcox_test(MFI ~ Particle_month, 
                                  paired = FALSE, 
                                  p.adjust.method = "BH", 
                                  data = x)
    
    # Round resulting p-values to 3 digits
    stats$adj.p.value <- round(stats$p.adj, 3)
    
    # Convert adjusted p values to significance symbols
    stats <- stats %>% mutate(adj.p.value =
                                case_when((adj.p.value >= 0.05) ~ NA,
                                          
                                          (adj.p.value < 0.05 &
                                             adj.p.value >= 0.01) ~ '*',
                                          
                                          (adj.p.value < 0.01 &
                                             adj.p.value >= 0.001) ~ '**',
                                          
                                          (adj.p.value < 0.001) ~ '***'))
    
    
    stats$test <- 'Wilcox'
    stats <- stats[, -c(6)] # Remove p.sigif to retain only adj.p.value
    stats
           
  }
  
}

# For each particle month
Month_stats <- lapply(Month_stats, function(x){
  
  # Subset each treatment
  stats <- Vehicle %>% subset(Treatment == x)
  
  # Calculate statistics
  stats <- Month_statistics(stats)

  stats$Treatment <- x
  
  if(all(is.na(stats[, 'adj.p.value'])) == TRUE){
         
         stats$pos_x_1 <- 0
         stats$pos_x_2 <- 0
         stats$pos_x <- 0
         stats$pos_y <- 0
         stats
         
  } else{
    
    stats <- na.omit(stats)
    
    stats <- stats %>% mutate(pos_x_1 =
                              case_when((group1 == 'December') ~ 1,
                                        (group1 == 'March') ~ 2,
                                        (group1 == 'June') ~ 3,
                                        (group1 == 'September') ~ 4))
    
    stats <- stats %>% mutate(pos_x_2 =
                                case_when((group2 == 'December') ~ 1,
                                          (group2 == 'March') ~ 2,
                                          (group2 == 'June') ~ 3, 
                                          (group2 == 'September') ~ 4))
    
    stats$pos_x <- (stats$pos_x_1 + stats$pos_x_2)/2
    
    stats$pos_y <- seq(from = 0.25,
                       to = 0.25 * nrow(stats), 
                       0.25)
    stats
    
  }
})

# Combine the statistics into one place
Month_stats <- do.call(rbind, Month_stats)

# Determine the maximum data point
max <- max(Vehicle$MFI)

# Define the locations for placement of significance bars
pos <- rep(c(max, max * 1.05, max * 1.1, max * 1.15, max * 1.2, max * 1.25 ), 4)

# Define the shapes of each data point
# Shapes are shared between biological replicates between treatment groups
shapes <- c(15, 15, 15, 16,
            15, 16, 16, 17, 18,
            17, 17, 18, 12, 12,
            16, 17, 18, 18)

# Generate Supplemental Figure S1B
Supplemental_Figure_S1B <- ggplot(data = Vehicle, 
                                  aes(x = Particle_month, 
                                      y = MFI, 
                                      fill = Treatment), 
                                  show.legend = FALSE) +
  
  # Create a bar for each particle month within each treatment group
  geom_bar(aes(col = Particle_month), 
           position = position_dodge(),
           fill = 'white', 
           stat = "summary", 
           fun = "mean", 
           show.legend = FALSE) +
  
  # Create individual points for each data point
  # Set data point shapes to shapes indicated above
  geom_point(aes(col = Particle_month, 
                 shape = Sample.ID), 
             stat = "identity",
             position = position_dodge(width = .9), 
             alpha = .8,
             size = 5, 
             show.legend = FALSE) + 
  
  # Add SEM bars
  stat_summary(geom = "errorbar", 
               fun.data = mean_se, 
               position = "dodge") +
  
  xlab("Treatment") +
  
  ylab("Vehicle S. aureus \n Bioparticle Phagocytosis \n (MFI)") +
  
  labs(fill = "Particle Month") +
  
  facet_wrap(~factor(Treatment,
                     c('M0_Veh', 'M2_Veh', 'M0_M1', 'M2_M1')),
             nrow = 1) +
  
  theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 26),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 30), 
        
        plot.title = element_text(face = 'bold', 
                                  size = 20),
        
        legend.text = element_text(size = 20),
        
        legend.title = element_text(size = 20),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  
  scale_color_manual(name = 'Particle Month', 
                     breaks = c('December', 'March', 'June', 'September'),
                     values = c('December' = 'blue', 'March' = 'red', 
                                'June' = 'darkgreen', 'September' = 'purple')) +
  
  scale_shape_manual(values = shapes)

###########

# Generate Supplemental Figure S1B statistics bars
Supplemental_Figure_S1B_Statistics <- ggplot(data = Month_stats, 
                                             aes(x = group1,
                                                 y = pos_y)) +
  
  geom_segment(data = na.omit(Month_stats),
               aes(x = group1,
                   xend = group2,
                   y = pos_y,
                   yend = pos_y),
               colour = "black",
               linewidth = 1.25) +
  
  geom_text(data = Month_stats,
            label = Month_stats$adj.p.value,
            x = Month_stats$pos_x,
            y = Month_stats$pos_y,
            size = 14,
            colour = "black") +
  
  ylab("Vehicle S. aureus \n Bioparticle Phagocytosis \n (MFI)") +
  
  labs(fill = "Particle Month") +
  
  facet_wrap(~factor(Treatment,
                     c('M0_Veh', 'M2_Veh', 'M0_M1', 'M2_M1')),
             nrow = 1) +
  
  theme(axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.text.x = element_blank(),
        
        axis.title.x = element_blank(),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  
  scale_x_discrete(limits = c('December', 'March', 'June', 'September')) +
  scale_y_continuous(limits = c(0, max(Month_stats$pos_y) + 0.25))

##########

# Combine Supplemental Figure S1B data and statistics plots
tiff('Supplemental_Figure_S1B.tiff', height = 1000, width = 1500)

patchwork::wrap_plots(Supplemental_Figure_S1B_Statistics +  
                        theme_void(), 
                      Supplemental_Figure_S1B + 
                        theme(strip.text.x = element_text(size = 25)), 
                      ncol = 1,
                      heights = c(0.4, 2)) %>% 
  ggdraw()

dev.off()
```

### Calculate the fold change phagocytosis MFI of PM exposed samples versus vehicle exposed samples

```{r}
# Isolate individual PM treatments and matched vehicle controls
M0_Veh <- Phagocytosis_Raw %>% select(all_of(c('M0_Veh', 'M0_PM')))
M2_Veh <- Phagocytosis_Raw %>% select(all_of(c('M2_Veh', 'M2_PM')))
M0_M1 <- Phagocytosis_Raw %>% select(all_of(c('M0_M1', 'M0_M1.PM')))
M2_M1 <- Phagocytosis_Raw %>% select(all_of(c('M2_M1', 'M2_M1.PM')))

list <- list(M0_Veh, M2_Veh, M0_M1, M2_M1)

# Calculate the fold change versus corresponding vehicle
FC <- lapply(list, function(x){
  
  tmp <- data.frame(x[, 2]/x[, 1])
  
})

# Combine all data to one location, set group names, and add sample information
FC <- do.call(cbind, FC)
names <- c('M0_PM', 'M2_PM', 'M0_M1.PM', 'M2_M1.PM')
colnames(FC) <- names
rownames(FC) <- rownames(Phagocytosis_Raw)
FC <- data.frame(Phagocytosis_Raw[1:2], FC)

# Set fold change data to long format
FC <- FC %>% 
  pivot_longer(cols = names, 
               names_to = 'Treatment',
               values_to = 'MFI')

FC$Sample.ID <- paste0(FC$Sample.ID, 
                       '_', 
                       FC$Particle_month)

FC$Treatment <- factor(FC$Treatment, 
                       levels = c('M0_PM', 'M2_PM',
                                  'M0_M1.PM', 'M2_M1.PM'))

FC$Particle_month <- factor(FC$Particle_month, 
                            levels = c('December', 'March', 'June', 'September'))

save(Phagocytosis = FC, 
     file = 'Fold_Change_Phagocytosis.RData')
```

### Calculate statistics and graph results for Figure 1A

```{r}
# Calculate statistics
stats <- statistics(FC)

# Determine the maximum data value for placement of significance bars
max <- max(FC$MFI)

# Define the locations for placement of significance bars
pos <- c(max, max * 1.05, max * 1.1, max * 1.15, max * 1.2, max * 1.25 )

# Generate Figure 1A
Figure_1A <- ggplot(data = FC, 
                    aes(x = Treatment, 
                        y = MFI)) + 
  
  # Create a bar for each treatment group mean
  geom_bar(aes(col = Treatment), 
           position = position_dodge(),
           fill = 'white', 
           stat = "summary", 
           fun = "mean", 
           show.legend = FALSE) +
  
  # Create individual points for each data point
  # Set data point shapes to shapes indicated above
  geom_jitter(aes(col = Treatment),
              width = 0.25,
              size = 4,
              show.legend = FALSE) + 
  
  # Add SEM bars
  stat_summary(aes(col = Treatment),
               geom = "errorbar", 
               fun.data = mean_se, 
               position = "dodge") +
  
  xlab("Treatment") +
  
  ylab("Fold Change Vehicle \n S. aureus Bioparticle Phagocytosis \n (MFI)") +
  
  labs(fill = "Particle Month") +
  
  theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 26),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 30),
        
        plot.title = element_text(face = 'bold', 
                                  size = 20),
        
        legend.text = element_text(size = 20),
        
        legend.title = element_text(size = 20),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  
  scale_color_manual(name = 'Treatment', 
                     breaks = c('M0_PM', 'M2_PM', 'M0_M1.PM', 'M2_M1.PM'),
                     values = c('M0_PM' = 'blue', 'M2_PM' = 'red', 
                                'M0_M1.PM' = 'darkgreen', 'M2_M1.PM' = 'purple'))

###########

# Generate Figure 1A statistics bars
Figure_1A_Statistics <- ggplot(data = stats, 
                               aes(x = Treatment,
                                   y = pos_y)) +
  
  geom_segment(data = na.omit(stats), 
               aes(x = group1, 
                   xend = group2,
                   y = pos_y, 
                   yend = pos_y), 
               colour = "black", 
               linewidth = 1.25) +
  
  geom_text(data = stats,
            label = stats$adj.p.value,
            x = stats$pos_x,
            y = stats$pos_y,
            size = 14,
            colour = "black") +
  
  ylab("Fold Change Vehicle \n S. aureus Biparticle Phagocytosis \n (MFI)") +
  
  labs(fill = "Particle Month") +
  
  theme(axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.text.x = element_blank(),

        axis.title.x = element_blank(),

        plot.margin = unit(c(0, 0, 0, 0), "cm")) +

  scale_x_discrete(limits = c('M0_PM', 'M2_PM',
                              'M0_M1.PM', 'M2_M1.PM')) +
  scale_y_continuous(limits = c(0, max(stats$pos_y) + 0.25))

##########

# Combine Figure 1A data and statistics plots
tiff('Figure_1A.tiff', height = 1000, width = 1500)

wrap_plots(Figure_1A_Statistics +  
             theme_void(),
           Figure_1A + 
             theme(legend.position = "none"), 
           ncol = 1,
           heights = c(0.4, 2)) %>% 
  ggdraw()

dev.off()
```

### Calculate statistics and graph results for Figure 1B

```{r}
Month_stats <- c('M0_PM', 'M2_PM', 'M0_M1.PM', 'M2_M1.PM')

# For each particle month
Month_stats <- lapply(Month_stats, function(x){
  
  # Subset each treatment
  stats <- FC %>% subset(Treatment == x)
  
  # Calculate statistics
  stats <- Month_statistics(stats)

  stats$Treatment <- x
  
  if(all(is.na(stats[, 'adj.p.value'])) == TRUE){
         
         stats$pos_x_1 <- 0
         stats$pos_x_2 <- 0
         stats$pos_x <- 0
         stats$pos_y <- 0
         stats
         
  } else{
    
    stats <- na.omit(stats)
    
    stats <- stats %>% mutate(pos_x_1 =
                              case_when((group1 == 'December') ~ 1,
                                        (group1 == 'March') ~ 2,
                                        (group1 == 'June') ~ 3,
                                        (group1 == 'September') ~ 4))
    
    stats <- stats %>% mutate(pos_x_2 =
                                case_when((group2 == 'December') ~ 1,
                                          (group2 == 'March') ~ 2,
                                          (group2 == 'June') ~ 3, 
                                          (group2 == 'September') ~ 4))
    
    stats$pos_x <- (stats$pos_x_1 + stats$pos_x_2)/2
    
    stats$pos_y <- seq(from = 0.25,
                       to = 0.25 * nrow(stats), 
                       0.25)
    stats
    
  }
})

# Combine the statistics into one place
Month_stats <- do.call(rbind, Month_stats)

# Determine the maximum data point
max <- max(FC$MFI)

# Define the locations for placement of significance bars
pos <- rep(c(max, max * 1.05, max * 1.1, max * 1.15, max * 1.2, max * 1.25 ), 4)

# Generate Supplemental Figure S1B
Figure_1B <- ggplot(data = FC, 
                    aes(x = Particle_month, 
                        y = MFI, 
                        fill = Treatment), 
                    show.legend = FALSE) + 
  
  # Create a bar for each particle month within each treatment group
  geom_bar(aes(col = Particle_month), 
           position = position_dodge(),
           fill = 'white', 
           stat = "summary", 
           fun = "mean", 
           show.legend = FALSE) +
  
  # Create individual points for each data point
  # Set data point shapes to shapes indicated above
  geom_point(aes(col = Particle_month, 
                 shape = Sample.ID), 
             stat = "identity",
             position = position_dodge(width = .9), 
             alpha = .8,
             size = 6, 
             show.legend = FALSE) + 

  # Add SEM bars
  stat_summary(geom = "errorbar", 
               fun.data = mean_se, 
               position = "dodge") +
  
  xlab("Treatment") +
  
  ylab("Fold Change Vehicle \n S. aureus Bioparticle Phagocytosis \n (MFI)") +
  
  labs(fill = "Particle Month") +
  
  facet_wrap(~factor(Treatment,
                     c('M0_PM', 'M2_PM', 'M0_M1.PM', 'M2_M1.PM')),
             nrow = 1) +

  theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 26),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 30),
        
        plot.title = element_text(face = 'bold', 
                                  size = 20),
        
        legend.text = element_text(size = 20),
        
        legend.title = element_text(size = 20),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  
  scale_color_manual(name = 'Particle Month', 
                     breaks = c('December', 'March', 'June', 'September'),
                     values = c('December' = 'blue', 'March' = 'red', 
                                'June' = 'darkgreen', 'September' = 'purple')) +
  
  scale_shape_manual(values = shapes)

###########

# Generate Figure 1B statistics bars
Figure_1B_Statistics <- ggplot(data = Month_stats, 
                               aes(x = Month,
                                   y = pos_y)) +
  
  geom_segment(data = na.omit(Month_stats),
               aes(x = group1,
                   xend = group2,
                   y = pos_y,
                   yend = pos_y),
               colour = "black",
               linewidth = 1.25) +
  
  geom_text(data = Month_stats,
            label = Month_stats$adj.p.value,
            x = Month_stats$pos_x,
            y = Month_stats$pos_y,
            size = 14,
            colour = "black") +
  
  ylab("Fold Change Vehicle \n S. aureus Bioparticle Phagocytosis \n (MFI)") +
  
  labs(fill = "Particle Month") +
  
  facet_wrap(~factor(Treatment,
                     c('M0_PM', 'M2_PM', 'M0_M1.PM', 'M2_M1.PM')),
             nrow = 1) +
  
  theme(axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.text.x = element_blank(),
        
        axis.title.x = element_blank(),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  
  scale_x_discrete(limits = c('December', 'March', 'June', 'September')) +
  scale_y_continuous(limits = c(0, max(Month_stats$pos_y) + 0.25))

##########

# Combine Figure 1B data and statistics plots
tiff('Figure_1B.tiff', height = 1000, width = 1500)

patchwork::wrap_plots(Figure_1B_Statistics +  
                        theme_void(), 
                      Figure_1B + 
                        theme(strip.text.x = element_text(size = 25)), 
                      ncol = 1,
                      heights = c(0.4, 2)) %>% 
  ggdraw()

dev.off()
```