---
title: "MSD Analysis"
author: "Timothy Smyth"
date: "2024-03-21"
output: html_document
---

# Analysis of MSD experiments

### Load packages 

```{r message = FALSE}
rm(list = ls(all.names = TRUE)) # clears global environ.

# Load packages
library(tidyverse) # for data cleaning
library(ggplot2) # for plotting
library(RColorBrewer) # for color palette
library(tibble)
library(tidyverse)
library(dplyr) 
library(parallelpam)
library(RColorBrewer)
library(ComplexHeatmap)
library(circlize)
library(digest)
library(cluster)
library(factoextra)
library(patchwork)
library(cowplot)
```

### Import and format data

```{r}
# Import data
MSD <- read.csv("MSD.csv")

colnames(MSD) <- c('Sample', 'Assay', 'Well', 'Spot', 
                   'Dilution', 'Concentration', 'Signal', 'Adjusted_signal', 
                   'Mean', 'Adj_signal_mean', 'CV', 'Percent_recovery', 
                   'Percent_Recovery_mean', 'Calc_conc', 'Calc_conc_mean', 'Calc_conc_CV')

# Remove Pro-inflammatory panel IL-8 as Chemokine IL-8 has much higher max concentration
MSD <- MSD %>% filter(Assay != "IL-8" | Spot != '6')

MSD$Month <- 'NA'

MSD <- MSD %>% dplyr::select(Month, everything())

# Remove unneeded variables
MSD <- MSD %>% subset(select = -c(Well, Spot, Dilution, Concentration, 
                                  Adjusted_signal, Mean, Signal, Adj_signal_mean, CV, 
                                  Percent_recovery, Percent_Recovery_mean, Calc_conc_mean, Calc_conc_CV))

# 289 and 290 shared between two PM groups. Those are already marked 
# with their month, so it is used below instead of the number
MSD <- MSD %>% mutate(Month =
                        case_when(str_detect(Sample, '191') |
                                    str_detect(Sample, '303') |  
                                    str_detect(Sample, 'Mar') |
                                    str_detect(Sample, '291') |
                                    str_detect(Sample, '132') ~ 'March', 
                                  
                                  str_detect(Sample, 'Jun') | 
                                    str_detect(Sample, '273') |
                                    str_detect(Sample, '280') |
                                    str_detect(Sample, '292') |
                                    str_detect(Sample, '288') ~ 'June', 
                                  
                                  str_detect(Sample, '312') |
                                    str_detect(Sample, '379') |
                                    str_detect(Sample, 'Sept') |
                                    str_detect(Sample, '308') |
                                    str_detect(Sample, '310') ~ 'September', 
                                  
                                  str_detect(Sample, 'Dec') |
                                    str_detect(Sample, '375') |
                                    str_detect(Sample, '285') |
                                    str_detect(Sample, '314') |
                                    str_detect(Sample, '388') ~ 'December'))

MSD$Group <- 'NA'
MSD <- MSD %>% dplyr::select(Group, everything())

# Rename groups
MSD <- MSD %>% mutate(Group = 
                        case_when(str_detect(Sample, 'M0_Veh') ~ 'M0_Veh', 
                                  str_detect(Sample, 'M2_Veh') ~ 'M2_Veh',
                                  
                                  str_detect(Sample, 'M0_M1') & str_detect(Sample, '\\+') ~ 'M0_M1+PM',
                                  str_detect(Sample, 'M2_M1') & str_detect(Sample, '\\+') ~ 'M2_M1+PM',
                                  
                                  str_detect(Sample, 'M0_M1') ~ 'M0_M1',
                                  str_detect(Sample, 'M2_M1') ~ 'M2_M1',
                                  
                                  str_detect(Sample, 'M0_PM') ~ 'M0_PM',
                                  str_detect(Sample, 'M2_PM') ~ 'M2_PM'))

MSD$Treatment <- 'NA'
MSD <- MSD %>% dplyr::select(Treatment, everything())

MSD <- MSD %>% mutate(Treatment = 
                        case_when(str_detect(Sample, 'Veh') ~ 'Veh', 
                                  str_detect(Sample, 'M1') & str_detect(Sample, '\\+') ~ 'M1+PM',
                                  str_detect(Sample, 'M1') ~ 'M1',
                                  str_detect(Sample, 'PM') ~ 'PM',
                        ))

MSD$Starting <- 'NA'
MSD <- MSD %>% dplyr::select(Starting, everything())

# Mark sample as starting as M0 or M2
MSD <- MSD %>% mutate(Starting = 
                        case_when(str_detect(Sample, 'M0') ~ 'M0', 
                                  str_detect(Sample, 'M2') ~ 'M2'))

# Set row names to sample and assay to prepare for pivot
MSD <- data.frame(MSD, row.names = paste0(MSD$Sample, "_", MSD$Assay))

# Pivot wider so each column is an analyte
MSD <- pivot_wider(MSD, names_from = Assay, values_from = Calc_conc)

# Set to data frame with row names as sample ID
MSD <- data.frame(MSD, row.names = MSD$Sample)

# Set up Month and Group column to ID experimental clustering below
MSD$Month_Group <- paste0(MSD$Month, "_", MSD$Group)
MSD <- MSD %>% dplyr::select(Month_Group, everything())

# Remove IFN-y and IL-4 due to presence in polarization/treatment media
MSD <- MSD %>% subset(select = -c(IFN.γ, IL.4))

# Remove IP-10 and MCP-4 since so many were above LOD
MSD <- MSD %>% subset(select = -c(IP.10, MCP.4))

MSD$Month <- factor(MSD$Month, 
                    levels = c('March', 
                               'June', 
                               'September', 
                               'December'))
```

### Generation of average cytokine/chemokine levels by treatment group - Figure 4A

```{r message = FALSE}
info <- MSD[1:6]

# Group samples by treatment group and calculate mean values
mean_MSD <- MSD[c(4, 7:ncol(MSD))] %>% group_by(Group) %>% summarize(across(everything(), mean))

# Scale data
heat <- scale(data.frame(mean_MSD[2:ncol(mean_MSD)], row.names = mean_MSD$Group))

# Perform PAM clustering
parallelpam::JWriteBin(as.matrix(t(heat)),
                       "MDM_mean.bin", 
                       dtype = "float", 
                       dmtype = "full")

CalcAndWriteDissimilarityMatrix("MDM_mean.bin", 
                                "MDM_mean2.bin",
                                distype = "L2", 
                                restype = "float",
                                comment = "L2 distance for vectors in jmatrix file vst_mean.bin", 
                                nthreads = 8)

JMatInfo("MDM_mean2.bin")

MDM_results_pam = ApplyPAM("MDM_mean2.bin", 
                           k = 4,
                           init_method = "BUILD", 
                           max_iter = 1000, 
                           nthreads = 8)

# Set color scheme for heatmap
myCol <- colorRamp2(c(-3, 0, 3), hcl_palette = "RdBu")

Group <- factor(unique(info$Group), 
                levels = c('M0_Veh',
                           'M2_Veh', 
                           'M0_PM',
                           'M2_PM',
                           'M0_M1',
                           'M2_M1',
                           'M0_M1+PM',
                           'M2_M1+PM'))

# Create annotation data frame
annotation <- data.frame(
  Group = Group,
  stringsAsFactors = FALSE)

colors <- list(
  
  Group = c('M0_Veh' = 'cornflowerblue',
            'M2_Veh' = 'forestgreen', 
            'M0_PM' = 'cadetblue2',
            'M2_PM' = 'darkseagreen3',
            'M0_M1' = 'tomato3',
            'M2_M1' = 'orange3',
            'M0_M1+PM' = 'tomato1',
            'M2_M1+PM' = 'orange1'))

colAnn <- HeatmapAnnotation(
  df = annotation,
  name = ' ',
  which = 'col', # 'col' (samples) or 'row' (gene) annotation?
  col = colors,
  annotation_height = 0.6,
  annotation_width = unit(1, 'cm'),
  gap = unit(1, 'mm'),
  show_legend = TRUE)

hmap <- Heatmap(as.matrix(t(heat)),
                
                 # split the genes / rows according to the PAM clusters
                 row_split = MDM_results_pam$clasif,
                 column_split = unique(info$Group),
                 cluster_row_slices = FALSE,
                 gap = unit(2.5, "mm"),
                 column_gap = unit(2.5, 'mm'),
                 border = TRUE,
                 height = ncol(heat) * unit(20, 'mm'),
                 width = ncol(heat) * unit(15, 'mm'),

                name = 'Z-Score',
                
                col = myCol,
                
                # Add value of each cell to the heatmap
                # Change the text color depending on the value of the cell
                cell_fun = function(j, i, x, y, width, height, fill) {
                  grid.text(sprintf("%.1f", heat[j, i]), x, y, gp = gpar(fontsize = 24, 
                                                                            col = if_else(heat[j, i] < 1.5 & 
                                                                                            heat[j, i] > -1.5 | 
                                                                                            is.na(heat[j, i]) == TRUE, 
                                                                                          'black', 
                                                                                          'white'),
                                                                            fontface = 'bold'))},
                
                # parameters for the color-bar that represents gradient of expression
                 heatmap_legend_param = list(
                   color_bar = 'continuous',
                   legend_direction = 'horizontal',
                   at = c(-3, -1.5, 0, 1.5, 3),
                   legend_width = unit(20, 'cm'),
                   legend_height = unit(30, 'cm'),
                   title_position = 'topcenter',
                   title_gp = gpar(fontsize = 30, fontface = 'bold'),
                   labels_gp = gpar(fontsize = 25, fontface = 'bold')),
                 
                 # row (gene) parameters
                 cluster_rows = TRUE,
                 show_row_dend = TRUE,
                 row_title_side = 'left',
                 row_title_gp = gpar(fontsize = 0,  fontface = 'bold'),
                 row_title_rot = 0,
                 show_row_names = TRUE,
                 row_names_gp = gpar(fontsize = 18, fontface = 'bold'),
                 row_names_side = 'left',
                 row_dend_width = unit(25,'mm'),
                 
                 # column (sample) parameters
                 cluster_columns = TRUE,
                 show_column_dend = TRUE,
                 column_title = NULL,
                 show_column_names = FALSE,
                 column_names_gp = gpar(fontsize = 10, fontface = 'bold'),
                 column_names_max_height = unit(10, 'cm'),
                 column_dend_height = unit(25,'mm'),
                 column_names_rot = 45,
                 
                 # cluster methods for rows and columns
                 clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
                 clustering_method_columns = 'ward.D2',
                 clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
                 clustering_method_rows = 'ward.D2',
                 
                 # specify top and bottom annotations
                 top_annotation = colAnn)
                 
png(file = "All_Groups_Heatmap.png", height = 3000, width = 2000)

draw(hmap,
     heatmap_legend_side = "bottom", 
     annotation_legend_side = "right", 
     legend_grouping = "original")

dev.off()
```

### Generation of PCA plot by treatment group - Figure 4B

```{r}
PCA <- data.frame(MSD[1:6], log(MSD[7:ncol(MSD)]))

col <- c('cornflowerblue',
         'forestgreen',
         'tomato3',
         'orange3', 
         'cadetblue2',
         'darkseagreen3',
         'tomato1',
         'orange1')


# Set value for shapes to be used below
shapes <- c(18, 18, 18, 18, 
            9, 9, 9, 9,
            15, 15, 15, 15,
            12, 12, 12, 12)

# Set groups to factor with following order
PCA$Group <- factor(PCA$Group,
                    levels = c('M0_Veh', 'M2_Veh', 
                               'M0_M1', 'M2_M1', 
                               'M0_PM', 'M2_PM',
                               'M0_M1+PM', 'M2_M1+PM'))

# Set groups to factor with following order
PCA$Month <- factor(PCA$Month, 
                    levels = c('December', 'March', 'June', 'September'))

#############################

fviz_pca_ind(prcomp(PCA[7:ncol(PCA)], center = TRUE, scale = TRUE),
             label = "var", 
             col.var = "grey50",
             habillage = PCA$Group, 
             palette = col,
             addEllipses = TRUE, 
             ellipse.type = 'confidence', 
             repel = TRUE, 
             ellipse.alpha = 0) + 
                
  ggtitle("PCA - Clustering") +
  theme(axis.title = element_text(size = 14), 
        axis.text = element_text(size = 12), 
        plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
        panel.border = element_rect(fill = NA, color = "black", size = 0.3),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlim(-10, 10) +
  
  # Add points for each group and color according to PM month
  geom_point(aes(color = as.factor(PCA$Group),
                 shape = as.factor(PCA$Group)),
             size = 3, 
             show.legend = FALSE) +
  
  # Manually set shapes to above indicated shapes
  scale_shape_manual(values = shapes) 
```

### Group fold changes

```{r}
# Break up everything into their subgroups
M0 <- MSD %>% subset(Group == 'M0_Veh' | Group == 'M0_PM')
M1 <- MSD %>% subset(Group == 'M0_M1' | Group == 'M0_M1+PM')
M2 <- MSD %>% subset(Group == 'M2_Veh' | Group == 'M2_PM')
M2_M1 <- MSD %>% subset(Group == 'M2_M1' | Group == 'M2_M1+PM')

# Compile into list for lapply
FC <- list(M0, M1, M2, M2_M1)

# Calculate the fold change of each treatment versus its related vehicle group
FC <- lapply(FC, function(x){
  
  # Vehicles are Veh or _M1
  # PM treatments are PM or M1+PM
  tmp1 <- x %>% subset(Treatment == 'Veh' | Treatment == 'M1') %>% subset(select = -c(1:6))
  tmp2 <- x %>% subset(Treatment == 'PM' | Treatment == 'M1+PM')
  info <- tmp2[1:6]
  tmp2 <- tmp2 %>% subset(select = -c(1:6))
  
  tmp3 <- tmp2/tmp1
  
  tmp3 <- cbind(info, tmp3)
  tmp3 <- tmp3 %>% pivot_longer(cols = colnames(tmp3[7:length(tmp3)]), 
                                names_to = 'Analyte', 
                                values_to = 'Value')
  
})

FC <- do.call(rbind, FC)

MSD <- MSD %>% pivot_longer(cols = colnames(MSD[7:length(MSD)]), 
                            names_to = 'Analyte', 
                            values_to = 'Raw_Value')

MSD <- MSD %>% subset(Treatment == 'PM' | Treatment == 'M1+PM')

merge <- merge(FC, MSD)

save(merge, 
     file = 'Fold_Change_MSD.RData')

merge <- merge %>% subset(Group == 'M0_M1+PM' |
                            Group == 'M2_M1+PM')

merge$Value <- log2(merge$Value)

merge$Group <- factor(merge$Group, 
                      levels = c('M0_Veh', 'M2_Veh', 
                                 'M0_M1', 'M2_M1', 
                                 'M0_PM', 'M2_PM',
                                 'M0_M1+PM', 'M2_M1+PM'))
                   
merge$Month <- factor(merge$Month, 
                      levels = c('December', 'March', 'June', 'September'))
```

### M1 fold change PCA Figure 5A

```{r}
M1_PCA <- FC %>% pivot_wider(names_from = 'Analyte', 
                             values_from = 'Value') %>% 
  subset(Group == 'M0_M1+PM'|
           Group == 'M2_M1+PM')

contrib <- fviz_contrib(prcomp(log(M1_PCA[7:ncol(M1_PCA)]), 
                               center = TRUE, 
                               scale = TRUE), 
                        choice = "var", 
                        axes = 1:2)

contrib <- contrib$data %>% 
  arrange(desc(contrib))

contrib <- contrib[1:10, ]

col <- c('cornflowerblue',
         'tomato3')

shapes <- c(15, 16, 17, 18, 
            0, 1, 2, 5, 
            6, 8)

# Set groups to factor with following order
M1_PCA$Group <- factor(M1_PCA$Group,
                       levels = c('M0_Veh', 'M2_Veh', 
                               'M0_M1', 'M2_M1', 
                               'M0_PM', 'M2_PM',
                               'M0_M1+PM', 'M2_M1+PM'))
                    
# Set groups to factor with following order
M1_PCA$Month <- factor(M1_PCA$Month, levels = c('December', 'March', 'June', 'September'))

#############################

fviz_pca_biplot(prcomp(log(M1_PCA[7:ncol(M1_PCA)]), center = TRUE, scale = TRUE),
                label = "var", 
                col.var = "grey50",
                habillage = M1_PCA$Group, 
                palette = col,
                addEllipses = TRUE, 
                ellipse.type = 'confidence', 
                repel = TRUE, 
                ellipse.alpha = 0) + 
  
  ggtitle("PCA - Clustering") +
  theme(axis.title = element_text(size = 14), 
        axis.text = element_text(size = 12), 
        plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
        panel.border = element_rect(fill = NA, color = "black", size = 0.3),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlim(-10, 10) +
  
  # Add points for each group and color according to PM month
  geom_point(aes(color = as.factor(M1_PCA$Group),
                 shape = as.factor(M1_PCA$Group)),
             size = 3, 
             show.legend = FALSE) +
  
  # Manually set shapes to above indicated shapes
  scale_shape_manual(values = shapes) 
```

### M1 fold change heatmap Figure 5B

```{r}
Heatmap <- FC %>% pivot_wider(names_from = 'Analyte', 
                              values_from = 'Value')

info <- unique(Heatmap[1:6] %>% subset(Group == 'M0_PM' | 
                                         Group == 'M2_PM')) %>% 
  arrange(Month) %>%
  mutate(Month = factor(Month, 
                        levels = c('December',
                                   'March',
                                   'June',
                                   'September')))

FC_mean_heat <- Heatmap %>% subset(Group == 'M0_M1+PM' | 
                                     Group == 'M2_M1+PM') %>% 
  group_by(Month_Group) %>% 
  summarize(across(everything(), mean))

FC_mean_heat <- data.frame(FC_mean_heat[7:length(FC_mean_heat)], 
                           row.names = FC_mean_heat$Month_Group)

FC_mean_heat <- FC_mean_heat[c(1, 5, 3, 7, 
                               2, 6, 4, 8), ]

FC_mean_heat <- t(scale(FC_mean_heat))

# Perform PAM clustering
parallelpam::JWriteBin(as.matrix(FC_mean_heat),
                       "MDM_mean.bin", 
                       dtype = "float", 
                       dmtype = "full")

CalcAndWriteDissimilarityMatrix("MDM_mean.bin", 
                                "MDM_mean2.bin",
                                distype = "L2", 
                                restype = "float",
                                comment = "L2 distance for vectors in jmatrix file vst_mean.bin", 
                                nthreads = 8)

JMatInfo("MDM_mean2.bin")

MDM_results_pam = ApplyPAM("MDM_mean2.bin", 
                           k = 4,
                           init_method = "BUILD", 
                           max_iter = 1000, 
                           nthreads = 8)

Group <- factor(str_sub(colnames(FC_mean_heat), 
                        -8, 
                        nchar(colnames(FC_mean_heat))), 
                levels = c('M0_M1+PM',
                           'M2_M1+PM'))

Month <- factor(str_sub(colnames(FC_mean_heat), 
                        1, 
                        -10), 
                levels = c('December',
                           'March',
                           'June',
                           'September'))

# Create annotation data frame
annotation <- data.frame(
  Group = Group, 
  Month = Month,
  stringsAsFactors = FALSE)

colors <- list(
  Group = c('M0_M1+PM' = 'cornflowerblue',
            'M2_M1+PM' = 'tomato3'), 
  
  Month = c('December' = 'blue',
            'March' = 'red',
            'June' = 'green',
            'September' = 'purple'))

colAnn <- HeatmapAnnotation(
  df = annotation,
  name = ' ',
  which = 'col', # 'col' (samples) or 'row' (gene) annotation?
  col = colors,
  annotation_height = 0.6,
  annotation_width = unit(1, 'cm'),
  gap = unit(1, 'mm'),
  show_legend = TRUE)

hmap2 <- Heatmap(FC_mean_heat,
                 
                 # split the genes / rows according to the PAM clusters
                 row_split = MDM_results_pam$clasif,
                 column_split = Group,
                 cluster_row_slices = FALSE,
                 gap = unit(2.5, "mm"),
                 column_gap = unit(2.5, 'mm'),
                 border = TRUE,
                 height = ncol(FC_mean_heat) * unit(20, 'mm'),
                 width = ncol(FC_mean_heat) * unit(15, 'mm'),
                 
                 name = 'Z-Score',

                # Add value of each cell to the heatmap
                # Change the text color depending on the value of the cell
                cell_fun = function(j, i, x, y, width, height, fill) {
                  grid.text(sprintf("%.1f", t(FC_mean_heat)[j, i]), x, y, gp = gpar(fontsize = 9, 
                                                                         col = 'black',
                                                                         fontface = 'bold'))},
                
                 col = myCol,
                 
                 # parameters for the color-bar that represents gradient of expression
                 heatmap_legend_param = list(
                   color_bar = 'continuous',
                   legend_direction = 'horizontal',
                   at = c(-3, -1.5, 0, 1.5, 3),
                   legend_width = unit(8, 'cm'),
                   legend_height = unit(5.0, 'cm'),
                   title_position = 'topcenter',
                   title_gp=gpar(fontsize = 12, fontface = 'bold'),
                   labels_gp=gpar(fontsize = 12, fontface = 'bold')),
                 
                 # row (gene) parameters
                 cluster_rows = TRUE,
                 show_row_dend = TRUE,
                 row_title_side = 'left',
                 row_title_gp = gpar(fontsize = 0,  fontface = 'bold'),
                 row_title_rot = 0,
                 show_row_names = TRUE,
                 row_names_gp = gpar(fontsize = 9, fontface = 'bold'),
                 row_names_side = 'left',
                 row_dend_width = unit(25,'mm'),
                 
                 # column (sample) parameters
                 cluster_columns = FALSE,
                 show_column_dend = FALSE,
                 column_title = '',
                 column_title_side = 'bottom',
                 column_title_gp = gpar(fontsize = 10, fontface = 'bold'),
                 show_column_names = FALSE,
                 column_names_gp = gpar(fontsize = 10, fontface = 'bold'),
                 column_names_max_height = unit(10, 'cm'),
                 column_dend_height = unit(25,'mm'),
                 
                 # cluster methods for rows and columns
                 clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
                 clustering_method_columns = 'ward.D2',
                 clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
                 clustering_method_rows = 'ward.D2',
                 
                 # # specify top and bottom annotations
                 top_annotation = colAnn)


png(file = "M1_Groups_Mean_Heatmap.png", height = 800, width = 1000)

draw(hmap2,
     heatmap_legend_side = 'top')

dev.off()
```

### M1 fold change statistics and graphing - Figure 5C-L

```{r}
list <- as.list(unique(as.character(contrib$name)))
list2 <- as.list(unique(merge$Group))

log_stats <- lapply(list, function(x){
  
  tmp <- merge %>% subset(Analyte == x) %>% 
    arrange(Value) %>%    
    mutate(Month = factor(Month, 
                          levels = c('December', 'March', 'June', 'September')))
  
  tmp <- na.omit(tmp)
  
  tmp2 <- lapply(list2, function(y){
    
    tmp2 <- tmp %>% subset(Group == y)
    
    test <- tmp2 %>%
      group_by(Month) %>% 
      do(broom::tidy(shapiro.test(.$Value)))
    
    if(min(test$p.value) >= 0.05){
      
      stats <- rstatix::pairwise_t_test(Value ~ Month, 
                                        paired = FALSE, 
                                        p.adjust.method = "BH", 
                                        data = tmp2)
      
      stats$adj.p.value <- round(stats$p.adj, 3)
      
      # Convert adjusted p values to significance symbols
      stats <- stats %>% mutate(adj.p.value = 
                                  case_when((adj.p.value >= 0.05) ~ NA,
                                            
                                            (adj.p.value < 0.05 &
                                               adj.p.value >= 0.01) ~ '*', 
                                            
                                            (adj.p.value < 0.01 &
                                               adj.p.value >= 0.001) ~ '**',
                                            
                                            (adj.p.value < 0.001) ~ '***'))
      
      stats$test <- 't.test'
      stats <- stats[, -c(7)] # Remove p.sigif to retain only adj.p.value
      stats$Treatment <- y
      stats
      
      } else{ # If treatment group does not have a normal distribution
        
        # Run a Wilcox test with BH correction
        stats <- rstatix::pairwise_wilcox_test(Value ~ Month, 
                                               paired = FALSE, 
                                               p.adjust.method = "BH", 
                                               data = tmp2)
        
        # Round resulting p-values to 3 digits
        stats$adj.p.value <- round(stats$p.adj, 3)
        
        # Convert adjusted p values to significance symbols
        stats <- stats %>% mutate(adj.p.value =
                                    case_when((adj.p.value >= 0.05) ~ NA,
                                              
                                              (adj.p.value < 0.05 &
                                                 adj.p.value >= 0.01) ~ '*',
                                              
                                              (adj.p.value < 0.01 &
                                                 adj.p.value >= 0.001) ~ '**',
                                              
                                              (adj.p.value < 0.001) ~ '***'))
        
        stats$test <- 'Wilcox'
        stats <- stats[, -c(6)] # Remove p.sigif to retain only adj.p.value
        stats$Treatment <- y
        stats
        
      }
    
    if(all(is.na(stats[, 'adj.p.value'])) == TRUE){
      
      stats$pos_x_1 <- 0
      stats$pos_x_2 <- 0
      stats$pos_x <- 0
      stats$pos_y <- 0
      stats
      
      } else{
        
        stats <- na.omit(stats)
        
        stats <- stats %>% mutate(pos_x_1 =
                                    case_when((group1 == 'December') ~ 1,
                                              (group1 == 'March') ~ 2,
                                              (group1 == 'June') ~ 3,
                                              (group1 == 'September') ~ 4))
        
        stats <- stats %>% mutate(pos_x_2 =
                                    case_when((group2 == 'December') ~ 1,
                                              (group2 == 'March') ~ 2,
                                              (group2 == 'June') ~ 3,
                                              (group2 == 'September') ~ 4))
        
        stats$pos_x <- (stats$pos_x_1 + stats$pos_x_2)/2
        
        stats$pos_y <- seq(from = 0.25,
                           to = 0.25 * nrow(stats), 
                           0.25)
        
        stats
      }
    
    })
  
  # Combine the statistics into one place
  tmp <- do.call(rbind, tmp2)
  tmp
  
  })

names(log_stats) <- list

# Define the shapes of each data point
# Shapes are shared between biological replicates between treatment groups
shapes <- c(15, 11, 15, 15, 16,
            15, 17, 16, 15, 15,
            16, 16, 16, 11, 17, 
            17, 18, 18, 16, 15,
            17, 16, 15, 11, 18,
            17, 11, 11, 11, 18, 
            17, 16, 18, 17, 11,
            18, 18, 17, 11, 18)

Figure_5_plots <- lapply(list, function(x){
  
  tmp <- merge %>% 
    subset(Analyte == x) %>% 
    arrange(Value) %>%    
    mutate(Month = factor(Month, 
                          levels = c('December', 'March', 'June', 'September')))
  
  tmp <- na.omit(tmp)
  
  max <- max(tmp$Value) * 1.05
  
  pos <- c(max, max * 1.15, max * 1.3, max * 1.45, max * 1.6, max * 1.75)
  
  ggplot(data = tmp, 
         aes(x = Month, 
             y = Value, 
             fill = Group)) +
    
    geom_bar(aes(col = Month),
             fill = 'white',
             position = position_dodge(),
             stat = "summary",
             fun = "mean",
             show.legend = FALSE) +
    
    geom_point(aes(col = Month,
                   shape = Sample),
               stat = "identity",
               position = position_dodge(width = .9),
               alpha = .8,
               size = 8,
               show.legend = FALSE) +
    
    # Add SEM bars
    stat_summary(geom = "errorbar",
                 fun.data = mean_se,
                 position = "dodge") +
    
    facet_wrap(~factor(Group, 
                       levels = c('M0_Veh', 'M2_Veh', 
                                  'M0_M1', 'M2_M1', 
                                  'M0_PM', 'M2_PM',
                                  'M0_M1+PM', 'M2_M1+PM')),
               nrow = 1) +
    
    # Label x and y axis
    labs(title = NULL,
         x = NULL, 
         y = paste0('Log2 Fold Change \n Vehicle ', x, ' Concentration'),
         fill = 'Particle Month') + 
    
    theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 30),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 32),
          
          strip.text.x = element_text(size = 32),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    
    scale_color_manual(name = 'Particle Month', 
                       breaks = c('December', 'March', 'June', 'September'),
                       values = c('December' = 'blue', 'March' = 'red', 
                                  'June' = 'darkgreen', 'September' = 'purple')) +
    
    scale_shape_manual(values = shapes)
  
})

names(Figure_5_plots) <- list

Figure_5_plots_stats <- lapply(list, function(x){
  
  tmp_plot_stats <- ggplot(data = log_stats[[x]], 
                           aes(x = group1,
                               y = pos_y)) +
    
    geom_segment(data = na.omit(log_stats[[x]]), 
                 aes(x = group1, 
                     xend = group2,
                     y = pos_y, 
                     yend = pos_y), 
                 colour = "black", 
                 linewidth = 1.25) +
    
    geom_text(data = log_stats[[x]],
              label = log_stats[[x]]$adj.p.value,
              x = log_stats[[x]]$pos_x,
              y = log_stats[[x]]$pos_y,
              size = 15,
              colour = "black") +
    
    facet_wrap(~factor(Treatment, 
                       levels = c('M0_Veh', 'M2_Veh', 
                                  'M0_M1', 'M2_M1', 
                                  'M0_PM', 'M2_PM',
                                  'M0_M1+PM', 'M2_M1+PM')),
               nrow = 1) +
    
    # Label x and y axis
    labs(title = NULL,
         x = NULL, 
         y = paste0('Log2 Fold Change \n Vehicle ', x, ' Concentration'),
         fill = 'Particle Month') + 
    
    theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 30),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 32),
          
          strip.text.x = element_text(size = 32),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    
    scale_x_discrete(limits = c('December', 'March', 'June', 'September')) +
    scale_y_continuous(limits = c(0, max(log_stats[[x]]$pos_y) + 0.25))
})

names(Figure_5_plots_stats) <- list

log_plot_combined <- list()

for(i in 1:length(list)){
  
  log_plot_combined[[i]] <- wrap_plots(Figure_5_plots_stats[[i]] +  
                                         theme_void(),
                                       Figure_5_plots[[i]] + 
                                         theme(legend.position = "none"), 
                                       ncol = 1,
                                       heights = c(0.4, 2)) %>% 
    ggdraw()
  
}
```

### M1 raw value statistics and graphing - Figure 5C-L

```{r}
list <- as.list(unique(as.character(contrib$name)))
list2 <- as.list(unique(merge$Group))

Raw_stats <- lapply(list, function(x){
  
  tmp <- merge %>% subset(Analyte == x) %>% 
    arrange(Raw_Value) %>%    
    mutate(Month = factor(Month, 
                          levels = c('December', 'March', 'June', 'September')))
  
  tmp <- na.omit(tmp)
  
  tmp2 <- lapply(list2, function(y){
    
    tmp2 <- tmp %>% subset(Group == y)
    
    test <- tmp2 %>%
      group_by(Month) %>% 
      do(broom::tidy(shapiro.test(.$Raw_Value)))
    
    if(min(test$p.value) >= 0.05){
      
      stats <- rstatix::pairwise_t_test(Raw_Value ~ Month,
                                        paired = FALSE, 
                                        p.adjust.method = "BH", 
                                        data = tmp2)
      
      stats$adj.p.value <- round(stats$p.adj, 3)
      
      # Convert adjusted p values to significance symbols
      stats <- stats %>% mutate(adj.p.value = 
                                  case_when((adj.p.value >= 0.05) ~ NA,
                                            
                                            (adj.p.value < 0.05 &
                                               adj.p.value >= 0.01) ~ '*', 
                                            
                                            (adj.p.value < 0.01 &
                                               adj.p.value >= 0.001) ~ '**',
                                            
                                            (adj.p.value < 0.001) ~ '***'))
      
      stats$test <- 't.test'
      stats <- stats[, -c(7)] # Remove p.sigif to retain only adj.p.value
      stats$Treatment <- y
      stats
      
      } else{ # If treatment group does not have a normal distribution
        
        # Run a Wilcox test with BH correction
        stats <- rstatix::pairwise_wilcox_test(Raw_Value ~ Month, 
                                               paired = FALSE, 
                                               p.adjust.method = "BH", 
                                               data = tmp2)
        
        # Round resulting p-values to 3 digits
        stats$adj.p.value <- round(stats$p.adj, 3)
        
        # Convert adjusted p values to significance symbols
        stats <- stats %>% mutate(adj.p.value =
                                    case_when((adj.p.value >= 0.05) ~ NA,
                                              
                                              (adj.p.value < 0.05 &
                                                 adj.p.value >= 0.01) ~ '*',
                                              
                                              (adj.p.value < 0.01 &
                                                 adj.p.value >= 0.001) ~ '**',
                                              
                                              (adj.p.value < 0.001) ~ '***'))
        
        stats$test <- 'Wilcox'
        stats <- stats[, -c(6)] # Remove p.sigif to retain only adj.p.value
        stats$Treatment <- y
        stats
        
      }
    
    if(all(is.na(stats[, 'adj.p.value'])) == TRUE){
      
      stats$pos_x_1 <- 0
      stats$pos_x_2 <- 0
      stats$pos_x <- 0
      stats$pos_y <- 0
      stats
      
      } else{
        
        stats <- na.omit(stats)
        
        stats <- stats %>% mutate(pos_x_1 =
                                    case_when((group1 == 'December') ~ 1,
                                              (group1 == 'March') ~ 2,
                                              (group1 == 'June') ~ 3,
                                              (group1 == 'September') ~ 4))
        
        stats <- stats %>% mutate(pos_x_2 =
                                    case_when((group2 == 'December') ~ 1,
                                              (group2 == 'March') ~ 2,
                                              (group2 == 'June') ~ 3,
                                              (group2 == 'September') ~ 4))
        
        stats$pos_x <- (stats$pos_x_1 + stats$pos_x_2)/2
        
        stats$pos_y <- seq(from = 0.25,
                           to = 0.25 * nrow(stats), 
                           0.25)
        
        stats
      }
    
    })
  
  # Combine the statistics into one place
  tmp <- do.call(rbind, tmp2)
  tmp
  
  })

names(Raw_stats) <- list

Figure_5_raw_plots <- lapply(list, function(x){
  
  tmp <- merge %>% 
    subset(Analyte == x) %>% 
    arrange(Raw_Value) %>%    
    mutate(Month = factor(Month, 
                          levels = c('December', 'March', 'June', 'September')))
  
  tmp <- na.omit(tmp)
  
  max <- max(tmp$Raw_Value) * 1.05
  
  pos <- c(max, max * 1.15, max * 1.3, max * 1.45, max * 1.6, max * 1.75)
  
  ggplot(data = tmp, 
         aes(x = Month, 
             y = Raw_Value, 
             fill = Group)) +
    
    geom_bar(aes(col = Month),
             fill = 'white',
             position = position_dodge(),
             stat = "summary",
             fun = "mean",
             show.legend = FALSE) +
    
    geom_point(aes(col = Month,
                   shape = Sample),
               stat = "identity",
               position = position_dodge(width = .9),
               alpha = .8,
               size = 8,
               show.legend = FALSE) +
    
    # Add SEM bars
    stat_summary(geom = "errorbar",
                 fun.data = mean_se,
                 position = "dodge") +
    
    facet_wrap(~factor(Group, 
                       levels = c('M0_Veh', 'M2_Veh', 
                                  'M0_M1', 'M2_M1', 
                                  'M0_PM', 'M2_PM',
                                  'M0_M1+PM', 'M2_M1+PM')),
               nrow = 1) +
    
    # Label x and y axis
    labs(title = NULL,
         x = NULL, 
         y = paste0(x, " Concentration \n (pg/mL)"), 
         fill = "Particle Month") + 
    
    theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 30),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 32),
          
          strip.text.x = element_text(size = 32),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    
    scale_color_manual(name = 'Particle Month', 
                       breaks = c('December', 'March', 'June', 'September'),
                       values = c('December' = 'blue', 'March' = 'red', 
                                  'June' = 'darkgreen', 'September' = 'purple')) +
    
    scale_shape_manual(values = shapes)
  
})

names(Figure_5_raw_plots) <- list

Figure_5_raw_plots_stats <- lapply(list, function(x){
  
  tmp_plot_stats <- ggplot(data = log_stats[[x]], 
                           aes(x = group1,
                               y = pos_y)) +
    
    geom_segment(data = na.omit(log_stats[[x]]), 
                 aes(x = group1, 
                     xend = group2,
                     y = pos_y, 
                     yend = pos_y), 
                 colour = "black", 
                 linewidth = 1.25) +
    
    geom_text(data = log_stats[[x]],
              label = log_stats[[x]]$adj.p.value,
              x = log_stats[[x]]$pos_x,
              y = log_stats[[x]]$pos_y,
              size = 15,
              colour = "black") +
    
    facet_wrap(~factor(Treatment, 
                       levels = c('M0_Veh', 'M2_Veh', 
                                  'M0_M1', 'M2_M1', 
                                  'M0_PM', 'M2_PM',
                                  'M0_M1+PM', 'M2_M1+PM')),
               nrow = 1) +
    
    # Label x and y axis
    labs(title = NULL,
         x = NULL, 
         y = paste0(x, " Concentration \n (pg/mL)"), 
         fill = 'Particle Month') + 
    
    theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 30),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 32),
          
          strip.text.x = element_text(size = 32),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    
    scale_x_discrete(limits = c('December', 'March', 'June', 'September')) +
    scale_y_continuous(limits = c(0, max(log_stats[[x]]$pos_y) + 0.25))
})

names(Figure_5_raw_plots_stats) <- list

raw_plots_combined <- list()

for(i in 1:length(list)){
  
  raw_plots_combined[[i]] <- wrap_plots(Figure_5_raw_plots_stats[[i]] +  
                                          theme_void(),
                                        Figure_5_raw_plots[[i]] + 
                                          theme(legend.position = "none"), 
                                        ncol = 1,
                                        heights = c(0.4, 2)) %>% 
    ggdraw()
  
}
```

### Combining M1 fold change and raw value graphs - Figure 5C-L

```{r}
for(i in 1:length(list)){
  
  tiff(paste0('Figure_5_', list[[i]], '.tiff'), width = 2000, height = 2000)
  
  print(
  
  wrap_plots(raw_plots_combined[[i]] + 
               theme_void(),
             log_plot_combined[[i]],
             ncol = 1,
             heights = c(1, 1)) %>% 
    ggdraw())
  
  dev.off()
}
```

### M0/M2 fold change PCA Figure 6A

```{r}
M0_M2_PCA <- FC %>% pivot_wider(names_from = 'Analyte', 
                             values_from = 'Value') %>% 
  subset(Group == 'M0_PM'|
           Group == 'M2_PM')

contrib <- fviz_contrib(prcomp(log(M0_M2_PCA[7:ncol(M0_M2_PCA)]), 
                               center = TRUE, 
                               scale = TRUE), 
                        choice = "var", 
                        axes = 1:2)

contrib <- contrib$data %>% 
  arrange(desc(contrib))

contrib <- contrib[1:10, ]

col <- c('forestgreen',
         'orange3')

shapes <- c(15, 16, 17, 18, 
            0, 1, 2, 5, 
            6, 8)

# Set groups to factor with following order
M0_M2_PCA$Group <- factor(M0_M2_PCA$Group,
                       levels = c('M0_Veh', 'M2_Veh', 
                               'M0_M0_M2', 'M2_M0_M2', 
                               'M0_PM', 'M2_PM',
                               'M0_M0_M2+PM', 'M2_M0_M2+PM'))
                    
# Set groups to factor with following order
M0_M2_PCA$Month <- factor(M0_M2_PCA$Month, levels = c('December', 'March', 'June', 'September'))

#############################

fviz_pca_biplot(prcomp(log(M0_M2_PCA[7:ncol(M0_M2_PCA)]), center = TRUE, scale = TRUE),
                label = "var", 
                col.var = "grey50",
                habillage = M0_M2_PCA$Group, 
                palette = col,
                addEllipses = TRUE, 
                ellipse.type = 'confidence', 
                repel = TRUE, 
                ellipse.alpha = 0) + 
  
  ggtitle("PCA - Clustering") +
  theme(axis.title = element_text(size = 14), 
        axis.text = element_text(size = 12), 
        plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
        panel.border = element_rect(fill = NA, color = "black", size = 0.3),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlim(-10, 10) +
  
  # Add points for each group and color according to PM month
  geom_point(aes(color = as.factor(M0_M2_PCA$Group),
                 shape = as.factor(M0_M2_PCA$Group)),
             size = 3, 
             show.legend = FALSE) +
  
  # Manually set shapes to above indicated shapes
  scale_shape_manual(values = shapes) 
```

### M0/M2 fold change heatmap Figure 6B

```{r}
Heatmap <- FC %>% pivot_wider(names_from = 'Analyte', 
                              values_from = 'Value')

info <- unique(Heatmap[1:6] %>% subset(Group == 'M0_PM' | 
                                         Group == 'M2_PM')) %>% 
  arrange(Month) %>%
  mutate(Month = factor(Month, 
                        levels = c('December',
                                   'March',
                                   'June',
                                   'September')))

FC_mean_heat <- Heatmap %>% subset(Group == 'M0_PM' | 
                                     Group == 'M2_PM') %>% 
  group_by(Month_Group) %>% 
  summarize(across(everything(), mean))

FC_mean_heat <- data.frame(FC_mean_heat[7:length(FC_mean_heat)], 
                           row.names = FC_mean_heat$Month_Group)

FC_mean_heat <- FC_mean_heat[c(1, 5, 3, 7, 
                               2, 6, 4, 8), ]

FC_mean_heat <- t(scale(FC_mean_heat))

# Perform PAM clustering
parallelpam::JWriteBin(as.matrix(FC_mean_heat),
                       "MDM_mean.bin", 
                       dtype = "float", 
                       dmtype = "full")

CalcAndWriteDissimilarityMatrix("MDM_mean.bin", 
                                "MDM_mean2.bin",
                                distype = "L2", 
                                restype = "float",
                                comment = "L2 distance for vectors in jmatrix file vst_mean.bin", 
                                nthreads = 8)

JMatInfo("MDM_mean2.bin")

MDM_results_pam = ApplyPAM("MDM_mean2.bin", 
                           k = 4,
                           init_method = "BUILD", 
                           max_iter = 1000, 
                           nthreads = 8)

Group <- factor(str_sub(colnames(FC_mean_heat), 
                        -5, 
                        nchar(colnames(FC_mean_heat))), 
                levels = c('M0_PM',
                           'M2_PM'))

Month <- factor(str_sub(colnames(FC_mean_heat), 
                        1, 
                        -7), 
                levels = c('December',
                           'March',
                           'June',
                           'September'))

# Create annotation data frame
annotation <- data.frame(
  Group = Group, 
  Month = Month,
  stringsAsFactors = FALSE)

colors <- list(
  Group = c('M0_PM' = 'forestgreen',
            'M2_PM' = 'orange3'), 
  
  Month = c('December' = 'blue',
            'March' = 'red',
            'June' = 'green',
            'September' = 'purple'))

colAnn <- HeatmapAnnotation(
  df = annotation,
  name = ' ',
  which = 'col', # 'col' (samples) or 'row' (gene) annotation?
  col = colors,
  annotation_height = 0.6,
  annotation_width = unit(1, 'cm'),
  gap = unit(1, 'mm'),
  show_legend = TRUE)

hmap3 <- Heatmap(FC_mean_heat,
                 
                 # split the genes / rows according to the PAM clusters
                 row_split = MDM_results_pam$clasif,
                 column_split = Group,
                 cluster_row_slices = FALSE,
                 gap = unit(2.5, "mm"),
                 column_gap = unit(2.5, 'mm'),
                 border = TRUE,
                 height = ncol(FC_mean_heat) * unit(20, 'mm'),
                 width = ncol(FC_mean_heat) * unit(15, 'mm'),
                 
                 name = 'Z-Score',

                # Add value of each cell to the heatmap
                # Change the text color depending on the value of the cell
                cell_fun = function(j, i, x, y, width, height, fill) {
                  grid.text(sprintf("%.1f", t(FC_mean_heat)[j, i]), x, y, gp = gpar(fontsize = 9, 
                                                                         col = 'black',
                                                                         fontface = 'bold'))},
                
                 col = myCol,
                 
                 # parameters for the color-bar that represents gradient of expression
                 heatmap_legend_param = list(
                   color_bar = 'continuous',
                   legend_direction = 'horizontal',
                   at = c(-3, -1.5, 0, 1.5, 3),
                   legend_width = unit(8, 'cm'),
                   legend_height = unit(5.0, 'cm'),
                   title_position = 'topcenter',
                   title_gp=gpar(fontsize = 12, fontface = 'bold'),
                   labels_gp=gpar(fontsize = 12, fontface = 'bold')),
                 
                 # row (gene) parameters
                 cluster_rows = TRUE,
                 show_row_dend = TRUE,
                 row_title_side = 'left',
                 row_title_gp = gpar(fontsize = 0,  fontface = 'bold'),
                 row_title_rot = 0,
                 show_row_names = TRUE,
                 row_names_gp = gpar(fontsize = 9, fontface = 'bold'),
                 row_names_side = 'left',
                 row_dend_width = unit(25,'mm'),
                 
                 # column (sample) parameters
                 cluster_columns = FALSE,
                 show_column_dend = FALSE,
                 column_title = '',
                 column_title_side = 'bottom',
                 column_title_gp = gpar(fontsize = 10, fontface = 'bold'),
                 show_column_names = FALSE,
                 column_names_gp = gpar(fontsize = 10, fontface = 'bold'),
                 column_names_max_height = unit(10, 'cm'),
                 column_dend_height = unit(25,'mm'),
                 
                 # cluster methods for rows and columns
                 clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
                 clustering_method_columns = 'ward.D2',
                 clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
                 clustering_method_rows = 'ward.D2',
                 
                 # # specify top and bottom annotations
                 top_annotation = colAnn)

png(file = "M0_M2_Groups_Mean_Heatmap.png", height = 800, width = 1000)

draw(hmap3,
     heatmap_legend_side = 'top')

dev.off()
```

### M0/M2 fold change statistics and graphing - Figure 6C-L

```{r}
merge <- merge(FC, MSD)

merge <- merge %>% subset(Group == 'M0_PM' |
                            Group == 'M2_PM')

merge$Value <- log2(merge$Value)

merge$Group <- factor(merge$Group, 
                      levels = c('M0_Veh', 'M2_Veh', 
                                 'M0_M1', 'M2_M1', 
                                 'M0_PM', 'M2_PM',
                                 'M0_M1+PM', 'M2_M1+PM'))
                   
merge$Month <- factor(merge$Month, 
                      levels = c('December', 'March', 'June', 'September'))

list <- as.list(unique(as.character(contrib$name)))
list2 <- as.list(unique(merge$Group))

log_stats <- lapply(list, function(x){
  
  tmp <- merge %>% subset(Analyte == x) %>% 
    arrange(Value) %>%    
    mutate(Month = factor(Month, 
                          levels = c('December', 'March', 'June', 'September')))
  
  tmp <- na.omit(tmp)
  
  tmp2 <- lapply(list2, function(y){
    
    tmp2 <- tmp %>% subset(Group == y)
    
    # TNF.b samples are all 0, and therefore cannot be tested for normality
    # Other months have non-normally distributed data (not shown), so Wilcox test
    # should be employed. Here, all other analytes are tested for normality and TNF.b
    # is directly tested with the Wilcox test
    if(x != 'TNF.β'){
      
      test <- tmp2 %>%
        group_by(Month) %>% 
        do(broom::tidy(shapiro.test(.$Value)))
      
      if(min(test$p.value) >= 0.05){
        
        stats <- rstatix::pairwise_t_test(Value ~ Month, 
                                          paired = FALSE,
                                          p.adjust.method = "BH", 
                                          data = tmp2)
        
        stats$adj.p.value <- round(stats$p.adj, 3)
        
        # Convert adjusted p values to significance symbols
        stats <- stats %>% mutate(adj.p.value = 
                                    case_when((adj.p.value >= 0.05) ~ NA,
                                              
                                              (adj.p.value < 0.05 &
                                                 adj.p.value >= 0.01) ~ '*', 
                                              
                                              (adj.p.value < 0.01 &
                                                 adj.p.value >= 0.001) ~ '**',
                                              
                                              (adj.p.value < 0.001) ~ '***'))
        
        stats$test <- 't.test'
        stats <- stats[, -c(7)] # Remove p.sigif to retain only adj.p.value
        stats$Treatment <- y
        stats
        
        } else{ # If treatment group does not have a normal distribution
          
          # Run a Wilcox test with BH correction
          stats <- rstatix::pairwise_wilcox_test(Value ~ Month, 
                                                 paired = FALSE,
                                                 p.adjust.method = "BH",
                                                 data = tmp2)
          
          # Round resulting p-values to 3 digits
          stats$adj.p.value <- round(stats$p.adj, 3)
          
          # Convert adjusted p values to significance symbols
          stats <- stats %>% mutate(adj.p.value =
                                      case_when((adj.p.value >= 0.05) ~ NA,
                                                
                                                (adj.p.value < 0.05 &
                                                   adj.p.value >= 0.01) ~ '*',
                                                
                                                (adj.p.value < 0.01 &
                                                   adj.p.value >= 0.001) ~ '**',
                                                
                                                (adj.p.value < 0.001) ~ '***'))
          
          stats$test <- 'Wilcox'
          stats <- stats[, -c(6)] # Remove p.sigif to retain only adj.p.value
          stats$Treatment <- y
          stats
          
        }
      } else{ 
        
        stats <- rstatix::pairwise_wilcox_test(Value ~ Month, 
                                               paired = FALSE, 
                                               p.adjust.method = "BH", 
                                               data = tmp2)
      
        stats$adj.p.value <- round(stats$p.adj, 3)
        
        stats <- stats %>% mutate(adj.p.value =
                                    case_when((adj.p.value >= 0.05) ~ NA,
                                              
                                              (adj.p.value < 0.05 &
                                                 adj.p.value >= 0.01) ~ '*',
                                              
                                              (adj.p.value < 0.01 &
                                                 adj.p.value >= 0.001) ~ '**',
                                              (adj.p.value < 0.001) ~ '***'))
        
        stats$Group <- y
        stats$test <- 'Wilcox'
        stats <- stats[, -c(6)]
        stats$Treatment <- y
        stats
        }
    
    if(all(is.na(stats[, 'adj.p.value'])) == TRUE){
      
      stats$pos_x_1 <- 0
      stats$pos_x_2 <- 0
      stats$pos_x <- 0
      stats$pos_y <- 0
      stats
      
      } else{
        
        stats <- na.omit(stats)

        stats <- stats %>% mutate(pos_x_1 =
                                    case_when((group1 == 'December') ~ 1,
                                              (group1 == 'March') ~ 2,
                                              (group1 == 'June') ~ 3,
                                              (group1 == 'September') ~ 4))
        
        stats <- stats %>% mutate(pos_x_2 =
                                    case_when((group2 == 'December') ~ 1,
                                              (group2 == 'March') ~ 2,
                                              (group2 == 'June') ~ 3,
                                              (group2 == 'September') ~ 4))
        
        stats$pos_x <- (stats$pos_x_1 + stats$pos_x_2)/2
        
        stats$pos_y <- seq(from = 0.25,
                           to = 0.25 * nrow(stats), 
                           0.25)
        
        stats
        
        }
    })
                
  tmp <- do.call(rbind, tmp2)

})

names(log_stats) <- list

shapes <- rep(c(15:18, 11), 8)

Figure_6_plots <- lapply(list, function(x){
  
  tmp <- merge %>% 
    subset(Analyte == x) %>% 
    arrange(Value) %>%    
    mutate(Month = factor(Month, 
                          levels = c('December', 'March', 'June', 'September')))
  
  tmp <- na.omit(tmp)
  
  max <- max(tmp$Value) * 1.05
  
  pos <- c(max, max * 1.15, max * 1.3, max * 1.45, max * 1.6, max * 1.75)
  
  ggplot(data = tmp, 
         aes(x = Month, 
             y = Value, 
             fill = Group)) +
    
    geom_bar(aes(col = Month),
             fill = 'white',
             position = position_dodge(),
             stat = "summary",
             fun = "mean",
             show.legend = FALSE) +
    
    geom_point(aes(col = Month,
                   shape = Sample),
               stat = "identity",
               position = position_dodge(width = .9),
               alpha = .8,
               size = 8,
               show.legend = FALSE) +
    
    # Add SEM bars
    stat_summary(geom = "errorbar",
                 fun.data = mean_se,
                 position = "dodge") +
    
    facet_wrap(~factor(Group, 
                       levels = c('M0_Veh', 'M2_Veh', 
                                  'M0_M1', 'M2_M1', 
                                  'M0_PM', 'M2_PM',
                                  'M0_M1+PM', 'M2_M1+PM')),
               nrow = 1) +
    
    # Label x and y axis
    labs(title = NULL,
         x = NULL, 
         y = paste0('Log2 Fold Change \n Vehicle ', x, ' Concentration'),
         fill = 'Particle Month') + 
    
    theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 30),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 32),
          
          strip.text.x = element_text(size = 32),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    
    scale_color_manual(name = 'Particle Month', 
                       breaks = c('December', 'March', 'June', 'September'),
                       values = c('December' = 'blue', 'March' = 'red', 
                                  'June' = 'darkgreen', 'September' = 'purple')) +
    
    scale_shape_manual(values = shapes)
  
})

names(Figure_6_plots) <- list

Figure_6_plots_stats <- lapply(list, function(x){
  
  tmp_plot_stats <- ggplot(data = log_stats[[x]], 
                           aes(x = group1,
                               y = pos_y)) +
    
    geom_segment(data = na.omit(log_stats[[x]]), 
                 aes(x = group1, 
                     xend = group2,
                     y = pos_y, 
                     yend = pos_y), 
                 colour = "black", 
                 linewidth = 1.25) +
    
    geom_text(data = log_stats[[x]],
              label = log_stats[[x]]$adj.p.value,
              x = log_stats[[x]]$pos_x,
              y = log_stats[[x]]$pos_y,
              size = 15,
              colour = "black") +
    
    facet_wrap(~factor(Treatment, 
                       levels = c('M0_Veh', 'M2_Veh', 
                                  'M0_M1', 'M2_M1', 
                                  'M0_PM', 'M2_PM',
                                  'M0_M1+PM', 'M2_M1+PM')),
               nrow = 1) +
    
    # Label x and y axis
    labs(title = NULL,
         x = NULL,
         y = paste0('Log2 Fold Change \n Vehicle ', x, ' Concentration'),
         fill = 'Particle Month') + 
    
    theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 30),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 32),
          
          strip.text.x = element_text(size = 32),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    
    scale_x_discrete(limits = c('December', 'March', 'June', 'September')) +
    scale_y_continuous(limits = c(0, max(log_stats[[x]]$pos_y) + 0.25))
})

names(Figure_6_plots_stats) <- list

log_plot_combined <- list()

for(i in 1:length(list)){
  
  log_plot_combined[[i]] <- wrap_plots(Figure_6_plots_stats[[i]] +  
                                         theme_void(),
                                       Figure_6_plots[[i]] + 
                                         theme(legend.position = "none"), 
                                       ncol = 1,
                                       heights = c(0.4, 2)) %>% 
    ggdraw()
  
}
```

### M0/M2 raw value statistics and graphing - Figure 6C-L

```{r}
list <- as.list(unique(as.character(contrib$name)))
list2 <- as.list(unique(merge$Group))

Raw_stats <- lapply(list, function(x){
  
  tmp <- merge %>% subset(Analyte == x) %>% 
    arrange(Raw_Value) %>%    
    mutate(Month = factor(Month, 
                          levels = c('December', 'March', 'June', 'September')))
  
  tmp <- na.omit(tmp)
  
  tmp2 <- lapply(list2, function(y){
    
    tmp2 <- tmp %>% subset(Group == y)
    
    # TNF.b samples are all 0, and therefore cannot be tested for normality
    # Other months have non-normally distributed data (not shown), so Wilcox test
    # should be employed. Here, all other analytes are tested for normality and TNF.b
    # is directly tested with the Wilcox test
    if(x != 'TNF.β'){
      
      test <- tmp2 %>%
        group_by(Month) %>% 
        do(broom::tidy(shapiro.test(.$Raw_Value)))
      
      if(min(test$p.value) >= 0.05){
        
        stats <- rstatix::pairwise_t_test(Raw_Value ~ Month, 
                                          paired = FALSE,
                                          p.adjust.method = "BH", 
                                          data = tmp2)
        
        stats$adj.p.value <- round(stats$p.adj, 3)
        
        # Convert adjusted p values to significance symbols
        stats <- stats %>% mutate(adj.p.value = 
                                    case_when((adj.p.value >= 0.05) ~ NA,
                                              
                                              (adj.p.value < 0.05 &
                                                 adj.p.value >= 0.01) ~ '*', 
                                              
                                              (adj.p.value < 0.01 &
                                                 adj.p.value >= 0.001) ~ '**',
                                              
                                              (adj.p.value < 0.001) ~ '***'))
        
        stats$test <- 't.test'
        stats <- stats[, -c(7)] # Remove p.sigif to retain only adj.p.value
        stats$Treatment <- y
        stats
        
        } else{ # If treatment group does not have a normal distribution
          
          # Run a Wilcox test with BH correction
          stats <- rstatix::pairwise_wilcox_test(Raw_Value ~ Month, 
                                                 paired = FALSE,
                                                 p.adjust.method = "BH",
                                                 data = tmp2)
          
          # Round resulting p-values to 3 digits
          stats$adj.p.value <- round(stats$p.adj, 3)
          
          # Convert adjusted p values to significance symbols
          stats <- stats %>% mutate(adj.p.value =
                                      case_when((adj.p.value >= 0.05) ~ NA,
                                                
                                                (adj.p.value < 0.05 &
                                                   adj.p.value >= 0.01) ~ '*',
                                                
                                                (adj.p.value < 0.01 &
                                                   adj.p.value >= 0.001) ~ '**',
                                                
                                                (adj.p.value < 0.001) ~ '***'))
          
          stats$test <- 'Wilcox'
          stats <- stats[, -c(6)] # Remove p.sigif to retain only adj.p.value
          stats$Treatment <- y
          stats
          
        }
      } else{ 
        
        stats <- rstatix::pairwise_wilcox_test(Raw_Value ~ Month, 
                                               paired = FALSE, 
                                               p.adjust.method = "BH", 
                                               data = tmp2)
      
        stats$adj.p.value <- round(stats$p.adj, 3)
        
        stats <- stats %>% mutate(adj.p.value =
                                    case_when((adj.p.value >= 0.05) ~ NA,
                                              
                                              (adj.p.value < 0.05 &
                                                 adj.p.value >= 0.01) ~ '*',
                                              
                                              (adj.p.value < 0.01 &
                                                 adj.p.value >= 0.001) ~ '**',
                                              (adj.p.value < 0.001) ~ '***'))
        
        stats$Group <- y
        stats$test <- 'Wilcox'
        stats <- stats[, -c(6)]
        stats$Treatment <- y
        stats
        }
    
    if(all(is.na(stats[, 'adj.p.value'])) == TRUE){
      
      stats$pos_x_1 <- 0
      stats$pos_x_2 <- 0
      stats$pos_x <- 0
      stats$pos_y <- 0
      stats
      
      } else{
        
        stats <- na.omit(stats)
        
        stats <- stats %>% mutate(pos_x_1 =
                                    case_when((group1 == 'December') ~ 1,
                                              (group1 == 'March') ~ 2,
                                              (group1 == 'June') ~ 3,
                                              (group1 == 'September') ~ 4))
        
        stats <- stats %>% mutate(pos_x_2 =
                                    case_when((group2 == 'December') ~ 1,
                                              (group2 == 'March') ~ 2,
                                              (group2 == 'June') ~ 3,
                                              (group2 == 'September') ~ 4))
        
        stats$pos_x <- (stats$pos_x_1 + stats$pos_x_2)/2
        
        stats$pos_y <- seq(from = 0.25,
                           to = 0.25 * nrow(stats), 
                           0.25)
        
        stats
        
        }
    })
                
  tmp <- do.call(rbind, tmp2)

})

names(Raw_stats) <- list

Figure_6_raw_plots <- lapply(list, function(x){
  
  tmp <- merge %>% 
    subset(Analyte == x) %>% 
    arrange(Raw_Value) %>%    
    mutate(Month = factor(Month, 
                          levels = c('December', 'March', 'June', 'September')))
  
  tmp <- na.omit(tmp)
  
  max <- max(tmp$Raw_Value) * 1.05
  
  pos <- c(max, max * 1.15, max * 1.3, max * 1.45, max * 1.6, max * 1.75)
  
  ggplot(data = tmp, 
         aes(x = Month, 
             y = Raw_Value, 
             fill = Group)) +
    
    geom_bar(aes(col = Month),
             fill = 'white',
             position = position_dodge(),
             stat = "summary",
             fun = "mean",
             show.legend = FALSE) +
    
    geom_point(aes(col = Month,
                   shape = Sample),
               stat = "identity",
               position = position_dodge(width = .9),
               alpha = .8,
               size = 8,
               show.legend = FALSE) +
    
    # Add SEM bars
    stat_summary(geom = "errorbar",
                 fun.data = mean_se,
                 position = "dodge") +
    
    facet_wrap(~factor(Group, 
                       levels = c('M0_Veh', 'M2_Veh', 
                                  'M0_M1', 'M2_M1', 
                                  'M0_PM', 'M2_PM',
                                  'M0_M1+PM', 'M2_M1+PM')),
               nrow = 1) +
    
    # Label x and y axis
    labs(title = NULL,
         x = NULL, 
         y = paste0(x, " Concentration \n (pg/mL)"), 
         fill = "Particle Month") + 
    
    theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 30),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 32),
          
          strip.text.x = element_text(size = 32),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    
    scale_color_manual(name = 'Particle Month', 
                       breaks = c('December', 'March', 'June', 'September'),
                       values = c('December' = 'blue', 'March' = 'red', 
                                  'June' = 'darkgreen', 'September' = 'purple')) +
    
    scale_shape_manual(values = shapes)
  
})

names(Figure_6_raw_plots) <- list

Figure_6_raw_plots_stats <- lapply(list, function(x){
  
  tmp_plot_stats <- ggplot(data = log_stats[[x]], 
                           aes(x = group1,
                               y = pos_y)) +
    
    geom_segment(data = na.omit(log_stats[[x]]), 
                 aes(x = group1, 
                     xend = group2,
                     y = pos_y, 
                     yend = pos_y), 
                 colour = "black", 
                 linewidth = 1.25) +
    
    geom_text(data = log_stats[[x]],
              label = log_stats[[x]]$adj.p.value,
              x = log_stats[[x]]$pos_x,
              y = log_stats[[x]]$pos_y,
              size = 15,
              colour = "black") +
    
    facet_wrap(~factor(Treatment, 
                       levels = c('M0_Veh', 'M2_Veh', 
                                  'M0_M1', 'M2_M1', 
                                  'M0_PM', 'M2_PM',
                                  'M0_M1+PM', 'M2_M1+PM')),
               nrow = 1) +
    
    # Label x and y axis
    labs(title = NULL,
         x = NULL, 
         y = paste0(x, " Concentration \n (pg/mL)"), 
         fill = 'Particle Month') + 
    
    theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1, 
                                   face="bold", 
                                   size = 30),
        
        axis.text.y = element_text(face="bold", 
                                   size = 26), 
        
        axis.title.x = element_text(face = 'bold', 
                                    size = 30), 
        
        axis.title.y = element_text(face = 'bold', 
                                    size = 32),
          
          strip.text.x = element_text(size = 32),
        
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    
    scale_x_discrete(limits = c('December', 'March', 'June', 'September')) +
    scale_y_continuous(limits = c(0, max(log_stats[[x]]$pos_y) + 0.25))
})

names(Figure_6_raw_plots_stats) <- list

raw_plots_combined <- list()

for(i in 1:length(list)){
  
  raw_plots_combined[[i]] <- wrap_plots(Figure_6_raw_plots_stats[[i]] +  
                                          theme_void(),
                                        Figure_6_raw_plots[[i]] + 
                                          theme(legend.position = "none"), 
                                        ncol = 1,
                                        heights = c(0.4, 2)) %>% 
    ggdraw()
  
}
```

### Combining M2 fold change and raw value graphs - Figure 6C-L

```{r}
for(i in 1:length(list)){
  
  tiff(paste0('Figure_6_', list[[i]], '.tiff'), width = 2000, height = 2000)
  
  print(
  
  wrap_plots(raw_plots_combined[[i]] + 
               theme_void(),
             log_plot_combined[[i]],
             ncol = 1,
             heights = c(1, 1)) %>% 
    ggdraw())
  
  dev.off()
}
```